\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,hyperref,array,xcolor,multicol,verbatim,mathpazo,algorithm,algpseudocode}
\usepackage[normalem]{ulem}
\usepackage[pdftex]{graphicx}

\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\begin{document}

\title{\vspace{-4cm}CS 270 Exam 1 True/False Repository}
\author{Neel Gupta}
\maketitle 

\begin{enumerate}
    \item Given a binary max-heap with n elements, the time complexity of find the smallest element is $O(\log n)$.
    \item A greedy algorithm considers the entire search space when making each step.
    \item Suppose that for a graph $G=(V,E)$ the average edge weight is $w$. Then a MST of $G$ will have weight at most ($V-1$)$w$.
    \item Consider two positively weighted graphs $G_1=(V,E,w_1)$ and $G_2=(V,E,w_2)$ with the same vertices and same edges s.t. every edge weighting in $G_2$ is the square of the edge weights in $G_1$. For any two vertices $u,v\in V$, any shortest path between $u$ and $v$ in $G_2$ is also a shortest path in $G_1$.
    \item If all edges in a connected undirected graph have unit cost, then you can find the MST using BFS.
    \item BFS is an example of a divide-and-conquer algorithm.
    \item For any cycle in a graph, the cheapest edge in the cycle is in a MST.
    \item DFS finds the longest paths from start vertex $s$ to each vertex $v$ in the graph.
    \item The shortest path in a weighted DAG can be found in linear time.
    \item Finding the $k$-th minimum element in an array of size $n$ using a binary min-heap takes $O(k\log n)$ time.
    \item We can merge any two arrays each of size $n$ into a new sorted array in $O(n)$.
    \item The shortest path in a weighted directed acyclic graph can be found in linear time.
    \item Given a weighted planar graph, Prim's algorithm using a binary heap implementation will outperform Prim's algorithm using an array implementation.
    \item If $f(n)=\Omega(n\log n)$ and $g(n)=O(n^2\log n)$, then $f(n)=O(g(n))$.
    \item Given a dense undirected weighted graph, the time for Prim's algorithm using a Fibonacci heap is $O(E)$.
    \item In a binomial min-heap with $n$ elements, the worst-case runtime complexity of finding the second smallest element is $O(1)$.
    \item Let $T$ be a complete binary tree with $n$ nodes. Finding a path from the root of $T$ to a given vertex $v\in T$ using BFS takes $O(\log n)$.
    \item Dijkstra's algorithm works correctly on a directed acyclic graph even when there are negative-weight edges.
    \item If the edge $e$ is not part of any MST of $G$, then it must be the maximum weight edge on some some cycle in $G$.
    \item If $f(n)=O(g(n))$ and $g(n)=O(f(n))$, then $f(n)=g(n)$.
    \item The following array is a binary max-heap: [10,3,5,1,4,2].
    \item There are at least 2 distinct solutions to the stable matching problem: one that is preferred by men and one that is preferrerd by women.
    \item In a binary max-heap with $n$ elements, the time complexity of finding the second largest element is $O(1)$.
    \item Given a binary max-heap with $n$ elements, the time complexity of find the smallest element is $O(\log n)$
    \item Kruskal's algorithm can fail in the presence of negative cost edges.
    \item If a weighted undirected graph has 2 MSTs, then its vertex set can be partioned into two, such that the minimum weight edge crossing the partition is not unique.
    \item In a connected undirected graph, and using the same starting point, the depth of any DFS tree is at least as much as the depth of any BFS tree.
    \item Algorithm A has runtime $\Theta(n^2)$ and algorithm B has runtime $\Theta(n\log n)$. From this we can conclude that A can never run faster than B on the same input set.
    \item Let $T$ be a complete binary tree with $n$ nodes. Finding a path from the root of $T$ to a given vertex $v\in T$ using BFS takes $O(\log n)$ time.
    \item Amortized cost of operations in a Fibonacci heap is at least as good as the worst case cost of those same operations in a binomial heap.
    \item Dijkstra's shortest path algorithm can be used to find shortest path in graphs with any edge weights.
    \item Function $f(n)=5n^2 4^n + 6n^4 3^n$ is $O(n^4 3^n)$.
    \item Consider stable matching. Suppose Jack preferse Rose to others, and Rose prefers Jack to others. The pair (Jack, Rose) exists in every stable matching.
    \item A DFS tree is a spanning tree.
    \item A binary max-heap can be built using an unsorted list of elements in $O(n)$ time.
    \item For some graphs BFS trees and DFS tress can be the same.
    \item The number of cycles in a bipartite graph may be odd.
    \item Stable matching algorithm presented in class is based on the greedy technique.
    \item To delete the $i$th node in a binary min-heap, you can exchange the last node with the $i$th node, then check the nodes below the $i$th node to see if the $i$th node should move down the heap to "re-heapify" it.
    \item If a connected undirected graph $G$ has the same weights for every edge, then a minimum spanning tree can be found in linear time.
    \item Given $n$ numbers, one could construct a binary heap using the $n$ numbers, then using the binary heap produce a sortest list of the numbers in $O(n)$ time.
    \item In a Fibonacci heap, the insert operations has an amoritzed cost of $O(1)$ time, but the worst case cost is higher.
    \item Function $10n^{10} 2^n + 3^n\log n$ is $O(n^{10} 2^n)$
    \item A directed graph has a topological ordering if and only if it contains no cycle.
    \item Both BFS and DFS can be used to find shortest path from one node to another node on graphs that are unweighted.
    \item If a directed graph has a topological ordering, then this topological ordering is unique.
    \item Kruskal's, Prim's, and Reverse-Delete algorithms are all examples of greedy algorithms.
    \item In an unweighted strongly connected (directed) graph, the shortest distance from A to B is always the same as that from B to A.
    \item In a weighted undirected graph, the shortest distance from A to B is always the same as that from B to A.
    \item In an unweighted directed graph, the shortest distance from A to B is always the same as that from B to A.
    \item In the stable matching problem with $n$ men and $n$ women, if a man and a woman are each other's last preferences, then they will never be matched with each other in any stable matching.
    \item We can find the $k$th largest element in a binary max-heap in $\Omega(1)$ time.
    \item A strongly connected (directed) graph cannot be a DAG.
    \item If the heaviest weight edge $e$ in an undirected connected graph $G$ is unique, then $e$ cannot belong to any MST of $G$.
    \item The height of a complete binary tree with $n$ nodes is $O(n)$.
    \item Given a graph $G$, if there is no negative cost cycles in $G$, then Dijkstra's algorithm will work correctly on $G$.
    \item In any graph, we have that $|E|=\Theta(|V|^2)$.
    \item If path p is the shortest path from $u$ to $v$ and $w$ is a node on the path, then the part of the path from $u$ to $w$ is also the shortest path from $u$ to $w$.
    \item Dijkstra's algorithm is able to find the shortest path in directed and undirected graphs with postitive edge weights.
    \item Nodes in a binomial heap can have more than 2 children.
    \item The following array is a binary max-heap: [16,14,10,10,12,9,3,2,4,1].
    \item In the stable matching problem involving $n$ men and $n$ women, for any given set of preference lists, there will be at most two stable matchings.
\end{enumerate}


\end{document}