\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,hyperref,array,xcolor,multicol,verbatim,mathpazo,algorithm,algpseudocode}
\usepackage[normalem]{ulem}
\usepackage[pdftex]{graphicx}

\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\begin{document}

\title{\vspace{-4cm}CS 270 Exam 1 Sample 1}
\author{Neel Gupta}
\maketitle 

\begin{problem}{1} 
    True/False Questions
\end{problem}
\textit{Answer:}
\begin{itemize}
    \item True. A cyclic directed graph cannot have a topological ordering.
    \item False. BFS can be used to find the shortest path on unweighted graphs.
    \item False. The same graph can have multiple topological orderings if sibling nodes aren't ordered.
    \item True. All MST algorithms are greedy because they always select an edge with a specific criteria to create the overall MST.
    \item False. A strongly connected graph means that there exists a path from all nodes to every other nodes, but since edges are directed, no bidirectionality means going the reverse direction can take much longer.
    \item True. The shortest path will be the same because edges are directed.
    \item True. The shortest path will be the same because edges are directed.
    \item False. Consider the following counter example. $M=(m_1,m_2)$ and $W=(w_1,w_2)$. If $m_1$ and $m_2$ both prefer $w_1$ over $w_2$ \textbf{and} $w_1$ and $w_2$ both prefer $m_1$ over $m_2$, then the following matching will result $\{(m_1,w_1),(m_2,w_2)\}$, where a man and a woman are each other's last preference and get matched.
    \item True. After every removal of the largest element, we need to trickle down which takes $\log n$ time, therefore $k \log n > c*1$ for some $c$.
    \item True. A strongly connected (directed) graph must have a cycle because if the paths from $u$ to $v$ and from $v$ to $u$ exist and are disjoint, then their union will create a cycle.
    \item True. The heaviest edge weight is removed until the graph becomes unconnected or acyclic in the Reverse Delete algorithm to create the MST.
    \item False. The height of a complete binary tree with $n$ nodes is $O(\log n)$.
    \item True. Dijkstra's will work on any non-negative cost edged graph.
\end{itemize}

\begin{problem}{2} 
    Multiple Choice Questions
\end{problem}
\textit{Answer:}
\begin{enumerate}
    \item (d). Finding an arbitrary element has a worst case $O(n)$ in a binary heap because there does not exist any ordering between left and right subtrees, and there is only an ordering between parent and children nodes.
    \item (a). Insertion/Bubble/Selection all run in $O(n^2)$, but they are lower bounded by $n^2$ as well since it does not do all $n$ iterations of work at each iteration. Rather it does some amount of work up to $n$ at each iteration.
    \item (c). Men end up with their best valid partner that they have not proposed to before.
    \item (a) \& (c). $\log n+n\log n$ and $n+\log(n^n)$ are both $\Theta(n\log n)$
    \item (a) \& (b). $n\log n$ and $n^2$ are bigger than $n$, but less than $n^3$.
    \item (e). The runtime of Dijkstra's algorithm using a binary heap is 
    $O(|E|\log|V|)$.
\end{enumerate}

\end{document}